Spatiotemporal Multi-Resolution Approximation of the Amari Type Neural Field Model

### 2012-01-17 Multiresolution_framework/ltx/manuscript.tex
working on MRA SS model section some more

Equations \ref{eq:KernelExpansion} and \ref{eq:FieldExpansion} are infinite series expansions and must be truncated to some level \dean{$J$} in order to solve the estimation problem. In other words, the neural field must be band-limited up to a given spatial frequency. Assuming the field and connectivity kernel are spatially band-limited they can be accurately approximated by
\begin{align}
	w\left(\mathbf{r}\right) &\approx \boldsymbol\theta^\top\boldsymbol\lambda\left(\mathbf{r}\right) 
	\label{eq:KernelFiniteExpansion} \\
	v_t\left(\mathbf{r}\right) &\approx \boldsymbol\mu^\top\left(\mathbf{r}\right)\mathbf{x}_t,
	\label{eq:FieldFiniteExpansion}
\end{align}
where the unknown parameter and state vectors, $\boldsymbol\theta \in \mathbb{R}^{n_{\theta}}$ and $\mathbf{x}_t \in \mathbb{R}^{n_x}$, are defined as 
\begin{align}
\boldsymbol\theta^\top &=[\begin{array}{ccccc} \boldsymbol\alpha_{j_0}^\top & \boldsymbol\beta_{j_0}^\top & \boldsymbol\beta_{j_0+1}^\top & \cdots & \boldsymbol\beta_{J}^\top \end{array}] 
\label{KernelWeights} \\
\mathbf{x}_{t}^\top &=[\begin{array}{ccccc}\mathbf{x}_{t,j_{0}}^\top &  \check{\mathbf{x}}_{t,j_{0}}^\top & \check{\mathbf{x}}_{t,j_{0}+1}^\top & \cdots & \check{\mathbf{x}}_{t,J}^\top\end{array}].
\label{FieldWeights}
\end{align}
The kernel and field approximation coefficient vectors, $\boldsymbol \alpha_{j_0}$ and $\mathbf{x}_{t,j_{0}}$, contain all the coefficients $\left\lbrace\alpha_{j_0, \mathbf{l}}:\mathbf{l} \in \mathbb{Z}^n \right\rbrace $ and $\left\lbrace x_{t,j_0, \mathbf{l}}: \mathbf{l} \in \mathbb{Z}^n\right\rbrace$, respectively. Similarly, the kernel and the field detail coefficient vectors, $\boldsymbol\beta_{j}$ and $\check{\mathbf{x}}_{t,j}$, contain all the coefficients $\left\lbrace \beta_{j,\mathbf{l}} :\mathbf{l} \in \mathbb{Z}^n\right\rbrace$ and $\left\lbrace  \check x_{t,j,\mathbf{l}}:\mathbf{l} \in \mathbb{Z}^n\right\rbrace$, respectively.

The vectors of the kernel and the field scaling and wavelet functions, $\boldsymbol\lambda$ and $\boldsymbol\mu$, respectively, are defined by
\begin{align}
    \label{KernelBasisVector}
    \boldsymbol\lambda^\top(\mathbf{r})=&\left[
    \begin{array}{ccccc}\boldsymbol\phi_{j_0}^\top(\mathbf{r}) &
    \boldsymbol\psi_{j_0}^\top(\mathbf{r}) & 
    \boldsymbol\psi_{j_0+1}^\top(\mathbf{r}) &
    \cdots &
    \boldsymbol\psi_{J}^\top(\mathbf{r})\end{array}\right] \\
    \label{FieldBasisVector}
    \boldsymbol\mu^\top (\mathbf{r}) =&\left[
    \begin{array}{ccccc}\boldsymbol\phi_{j_0}^\top(\mathbf{r}) &
    \boldsymbol\psi_{j_0}^\top(\mathbf{r}) & 
    \boldsymbol\psi_{j_0+1}^\top(\mathbf{r}) &
    \cdots &
    \boldsymbol\psi_{J}^\top(\mathbf{r})\end{array}\right]
\end{align}

The vectors in  (\ref{KernelBasisVector}) and (\ref{FieldBasisVector}) are constructed in the same manner as the vectors in  (\ref{KernelWeights}) and (\ref{FieldWeights}).

To derive the state-space model, we first define the matrices 
\begin{align}\label{eq:Lambdax}
 \mathbf{\Lambda}_{x} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right)\boldsymbol\mu^\top\left(r\right) dr \\
\label{eq:Lambdatheta}
 \mathbf{\Lambda}_{\theta} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right) \int_\Omega { 
	   \boldsymbol\theta^\top\boldsymbol\lambda\left(r-r'\right)
	    \boldsymbol\mu^\top\left(r\right)\ dr'dr.}
\end{align}
Substituting \eqref{eq:KernelFiniteExpansion} and \eqref{eq:FieldFiniteExpansion} in \eqref{eq:DiscreteTimeModel}, pre-multiplying by $\boldsymbol\mu\left(r\right)$, and integrating over the space gives
\begin{align}\label{eq:DecomposedModel2} 
	&\mathbf{\Lambda}_{x} \mathbf{x}_{t+1}= 
	\xi\mathbf{\Lambda}_{x} \mathbf{x}_t +T_s \varsigma \mathbf{\Lambda}_{\theta}\mathbf{x}_t +\int_{\Omega}\boldsymbol\mu\left(r\right)e_t\left(r\right)dr.
\end{align}
Cross-multiplying by $\mathbf{\Lambda}_{x}^{-1}$ gives the state transition equation
\begin{align}\label{eq:StateEquation}
 \mathbf x_{t+1} &=\mathbf A(\boldsymbol \theta) \mathbf x_t+ \mathbf w_t\\
\label{eq:A_theta}
 \mathbf A(\boldsymbol \theta) &= T_s\varsigma\mathbf{\Lambda}_{x}^{-1}\mathbf{\Lambda}_{\theta}+\xi\mathbf I.
\end{align}
The disturbance becomes 
\begin{equation}\label{eq:Disturbance}
\mathbf w_t= \mathbf{\Lambda}_{x}^{-1}\int_{\Omega}\boldsymbol\mu \left(r\right)e_t\left(r\right)dr,
\end{equation}
which is a vector valued, zero-mean normally distributed, white noise process with covariance (see \cite{Freestone2011} for proof)
\begin{equation}\label{eq:CovMatrix}
\boldsymbol\Sigma_w =\mathbf{\Lambda}_{x}^{-1}\iint\limits_{\boldsymbol\Omega}\boldsymbol\mu\left(r\right) \eta\left(r-r'\right)\boldsymbol\mu^{\top}\left(r'\right)dr'dr\mathbf{\Lambda}_{x}^{-\top}.
\end{equation}
The observation equation of the state-space model is found by substituting decomposition \eqref{eq:FieldFiniteExpansion}
 into \eqref{eq:ObservationEquation} giving
\begin{equation}\label{eq:ReducedObservationEquation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{equation}
where each element of the observation matrix is given by
\begin{equation}\label{eq:Observationmatrix}
	\mathbf{C}_{ij} \triangleq \int_{\Omega}m(r_i - r')\boldsymbol{\mu}_j(r') \, d\mathbf{r}'.
\end{equation}
\section{B-spline MRAIDE in State-Space }
B-spline functions are one of the most basic elements for wavelet construction, these functions are composed of  piecewise polynomials combined smoothly at some breaking points (knots) which are equally spaced in the case of cardinal B-splines. The degree of smoothness is determined by the order of splines \cite{Goswami1999} where the higher the order, the smoother the resulting B-spline becomes (see Figure~\ref{fig:Figure0}).
\begin{figure}[!ht]
\centering
\includegraphics{./Graph/Figure0.pdf}
% \includegraphics{./Graph/Figure0b.pdf}
\caption{{\bf Examples of cardinal B-spline functions}. Piecewise polynomial functions (coloured lines) are joined at the break points (black dots). (\textbf{A}) B-spline function of order 2 (linear B-spline). (\textbf{B}) B-spline function of order 8.}
\label{fig:Figure0}
\end{figure}

>>>>

Equations \ref{eq:KernelExpansion} and \ref{eq:FieldExpansion} are infinite series expansions and must be truncated to some level \dean{$j=J$} in order to solve the estimation problem. In other words, the neural field must be band-limited up to a given spatial frequency.

Now we will simplify the notation by defining vectors that contain all the translations of the wavelets and scaling functions such that
\begin{align}
	\boldsymbol\alpha_{k} \triangleq& \left\lbrace\alpha_{k, \mathbf{l}}:\mathbf{l} \in \mathbb{Z}^n \right\rbrace \\
	\boldsymbol\beta_{k} \triangleq& \\
	\mathbf{x}_{t,k} \triangleq& \\
	\check{\mathbf{x}}_{t,k} \triangleq& \\
	\boldsymbol\phi_{j_0}^\top(\mathbf{r}) \triangleq& \\
	\boldsymbol\psi_{k}^\top(\mathbf{r}) \triangleq& ,
\end{align}
where $k = j_0,j$.\dean{ARE THESE EACH $n_{\theta}$ or $n_x$ DIMENSIONAL VECTORS?} Following this we can define the \dean{vectors or matrices} 
\begin{align}
\boldsymbol\theta^\top \triangleq& [\begin{array}{ccccc} \boldsymbol\alpha_{j_0}^\top & \boldsymbol\beta_{j_0}^\top & \boldsymbol\beta_{j_0+1}^\top & \cdots & \boldsymbol\beta_{J}^\top \end{array}] 
\label{KernelWeights} \\
\mathbf{x}_{t}^\top \triangleq& [\begin{array}{ccccc}\mathbf{x}_{t,j_{0}}^\top &  \check{\mathbf{x}}_{t,j_{0}}^\top & \check{\mathbf{x}}_{t,j_{0}+1}^\top & \cdots & \check{\mathbf{x}}_{t,J}^\top\end{array}]
\label{FieldWeights} \\
\label{KernelBasisVector}
\boldsymbol\lambda^\top(\mathbf{r}) \triangleq& \left[
\begin{array}{ccccc} \boldsymbol\phi_{j_0}^\top(\mathbf{r}) &
\boldsymbol\psi_{j_0}^\top(\mathbf{r}) & 
\boldsymbol\psi_{j_0+1}^\top(\mathbf{r}) &
\cdots &
\boldsymbol\psi_{J}^\top(\mathbf{r})\end{array}\right] \\
\label{FieldBasisVector}
\boldsymbol\mu^\top (\mathbf{r}) \triangleq& \left[
\begin{array}{ccccc}\boldsymbol\phi_{j_0}^\top(\mathbf{r}) &
\boldsymbol\psi_{j_0}^\top(\mathbf{r}) & 
\boldsymbol\psi_{j_0+1}^\top(\mathbf{r}) &
\cdots &
\boldsymbol\psi_{J}^\top(\mathbf{r})\end{array}\right],
\end{align}
which allow us to write the neural field and connectivity kernel decomposition in the compact form of
\begin{align}
	w\left(\mathbf{r}\right) &\approx \boldsymbol\theta^\top\boldsymbol\lambda\left(\mathbf{r}\right) 
	\label{eq:KernelFiniteExpansion} \\
	v_t\left(\mathbf{r}\right) &\approx \boldsymbol\mu^\top\left(\mathbf{r}\right)\mathbf{x}_t.
	\label{eq:FieldFiniteExpansion}
\end{align}

% where the unknown parameter and state vectors, $\boldsymbol\theta \in \mathbb{R}^{n_{\theta}}$ and $\mathbf{x}_t \in \mathbb{R}^{n_x}$, \dean{($\boldsymbol\theta \in \mathbb{R}^{n_{\theta}\times (J+2)}$ and $\mathbf{x}_t \in \mathbb{R}^{n_x \times (J+2)}$)} are defined as 
% \begin{align}
% \boldsymbol\theta^\top \triangleq& [\begin{array}{ccccc} \boldsymbol\alpha_{j_0}^\top & \boldsymbol\beta_{j_0}^\top & \boldsymbol\beta_{j_0+1}^\top & \cdots & \boldsymbol\beta_{J}^\top \end{array}] 
% \label{KernelWeights} \\
% \mathbf{x}_{t}^\top \triangleq& [\begin{array}{ccccc}\mathbf{x}_{t,j_{0}}^\top &  \check{\mathbf{x}}_{t,j_{0}}^\top & \check{\mathbf{x}}_{t,j_{0}+1}^\top & \cdots & \check{\mathbf{x}}_{t,J}^\top\end{array}].
% \label{FieldWeights}
% \end{align}
% The kernel and field approximation coefficient vectors, $\boldsymbol \alpha_{j_0}$ and $\mathbf{x}_{t,j_{0}}$, contain all the coefficients $\left\lbrace\alpha_{j_0, \mathbf{l}}:\mathbf{l} \in \mathbb{Z}^n \right\rbrace $ and $\left\lbrace x_{t,j_0, \mathbf{l}}: \mathbf{l} \in \mathbb{Z}^n\right\rbrace$, respectively. Similarly, the kernel and the field detail coefficient vectors, $\boldsymbol\beta_{j}$ and $\check{\mathbf{x}}_{t,j}$, contain all the coefficients $\left\lbrace \beta_{j,\mathbf{l}} :\mathbf{l} \in \mathbb{Z}^n\right\rbrace$ and $\left\lbrace  \check x_{t,j,\mathbf{l}}:\mathbf{l} \in \mathbb{Z}^n\right\rbrace$, respectively.
% 
% The vectors of the kernel and the field scaling and wavelet functions, $\boldsymbol\lambda$ and $\boldsymbol\mu$, respectively, are defined by
% \begin{align}
%     \label{KernelBasisVector}
%     \boldsymbol\lambda^\top(\mathbf{r}) \triangleq& \left[
%     \begin{array}{ccccc} \boldsymbol\phi_{j_0}^\top(\mathbf{r}) &
%     \boldsymbol\psi_{j_0}^\top(\mathbf{r}) & 
%     \boldsymbol\psi_{j_0+1}^\top(\mathbf{r}) &
%     \cdots &
%     \boldsymbol\psi_{J}^\top(\mathbf{r})\end{array}\right] \\
%     \label{FieldBasisVector}
%     \boldsymbol\mu^\top (\mathbf{r}) \triangleq& \left[
%     \begin{array}{ccccc}\boldsymbol\phi_{j_0}^\top(\mathbf{r}) &
%     \boldsymbol\psi_{j_0}^\top(\mathbf{r}) & 
%     \boldsymbol\psi_{j_0+1}^\top(\mathbf{r}) &
%     \cdots &
%     \boldsymbol\psi_{J}^\top(\mathbf{r})\end{array}\right]
% \end{align}
% 
% The vectors in  (\ref{KernelBasisVector}) and (\ref{FieldBasisVector}) are constructed in the same manner as the vectors in  (\ref{KernelWeights}) and (\ref{FieldWeights}).  
Now we have the compact for of the decomposition we can substitute this into the neural field equation \eqref{eq:DiscreteTimeModel} giving
\begin{equation}\label{eq:DiscreteTimeModel}
	\boldsymbol\mu^\top\left(\mathbf{r}\right)\mathbf{x}_t = 
	\xi \boldsymbol\mu^\top\left(\mathbf{r}\right)\mathbf{x}_t + 
	T_s \int_\Omega { 
	    \boldsymbol\theta^\top\boldsymbol\lambda\left(\mathbf{r}-\mathbf{r}'\right)
	    f\left(\boldsymbol\mu^\top\left(\mathbf{r}'\right)\mathbf{x}_t\right) 
	\, \mathrm{d}\mathbf{r}'}  
	+ e_t\left(\mathbf{r}\right), 
\end{equation}

To derive the state-space model, we first define the matrices 
\begin{align}\label{eq:Lambdax}
 \mathbf{\Lambda}_{x} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right)\boldsymbol\mu^\top\left(r\right) dr \\
\label{eq:Lambdatheta}
 \mathbf{\Lambda}_{\theta} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right) \int_\Omega { 
	   \boldsymbol\theta^\top\boldsymbol\lambda\left(r-r'\right)
	    \boldsymbol\mu^\top\left(r\right)\ dr'dr.}
\end{align}
Substituting \eqref{eq:KernelFiniteExpansion} and \eqref{eq:FieldFiniteExpansion} in \eqref{eq:DiscreteTimeModel}, pre-multiplying by $\boldsymbol\mu\left(r\right)$, and integrating over the space gives
\begin{align}\label{eq:DecomposedModel2} 
	&\mathbf{\Lambda}_{x} \mathbf{x}_{t+1} = 
	\xi\mathbf{\Lambda}_{x} \mathbf{x}_t +T_s \varsigma \mathbf{\Lambda}_{\theta}\mathbf{x}_t +\int_{\Omega}\boldsymbol\mu\left(r\right)e_t\left(r\right)dr.
\end{align}
Cross-multiplying by $\mathbf{\Lambda}_{x}^{-1}$ gives the state transition equation
\begin{align}\label{eq:StateEquation}
 \mathbf x_{t+1} &=\mathbf A(\boldsymbol \theta) \mathbf x_t+ \mathbf w_t\\
\label{eq:A_theta}
 \mathbf A(\boldsymbol \theta) &= T_s\varsigma\mathbf{\Lambda}_{x}^{-1}\mathbf{\Lambda}_{\theta}+\xi\mathbf I.
\end{align}
The disturbance becomes 
\begin{equation}\label{eq:Disturbance}
\mathbf w_t= \mathbf{\Lambda}_{x}^{-1}\int_{\Omega}\boldsymbol\mu \left(r\right)e_t\left(r\right)dr,
\end{equation}
which is a vector valued, zero-mean normally distributed, white noise process with covariance (see \cite{Freestone2011} for proof)
\begin{equation}\label{eq:CovMatrix}
\boldsymbol\Sigma_w =\mathbf{\Lambda}_{x}^{-1}\iint\limits_{\boldsymbol\Omega}\boldsymbol\mu\left(r\right) \eta\left(r-r'\right)\boldsymbol\mu^{\top}\left(r'\right)dr'dr\mathbf{\Lambda}_{x}^{-\top}.
\end{equation}
The observation equation of the state-space model is found by substituting decomposition \eqref{eq:FieldFiniteExpansion}
 into \eqref{eq:ObservationEquation} giving
\begin{equation}\label{eq:ReducedObservationEquation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{equation}
where each element of the observation matrix is given by
\begin{equation}\label{eq:Observationmatrix}
	\mathbf{C}_{ij} \triangleq \int_{\Omega}m(r_i - r')\boldsymbol{\mu}_j(r') \, d\mathbf{r}'.
\end{equation}
\section{B-spline MRAIDE in State-Space }
B-spline functions are one of the most basic elements for wavelet construction, these functions are composed of  piecewise polynomials combined smoothly at some breaking points (knots) which are equally spaced in the case of cardinal B-splines. The degree of smoothness is determined by the order of splines \cite{Goswami1999} where the higher the order, the smoother the resulting B-spline becomes (see Figure~\ref{fig:Figure0}).
\begin{figure}[!ht]
\centering
\includegraphics{./Graph/Figure0.pdf}
% \includegraphics{./Graph/Figure0b.pdf}
\caption{{\bf Examples of cardinal B-spline functions}. Piecewise polynomial functions (coloured lines) are joined at the break points (black dots). (\textbf{A}) B-spline function of order 2 (linear B-spline). (\textbf{B}) B-spline function of order 8.}
\label{fig:Figure0}
\end{figure}
